---
title: "STATS506_ProbSet1"
author: "Heleyna Tucker"
format: html
editor: visual
---

# Problem 1

### Part A: Import the data into a 'data.frame'  in R. I have given the appropriate column name to the data frame as well: 

```{r}
col_names = c('class','Alcohol', 'Malic acid', 'Ash', 'Alcalinity of ash', 'Magnesium', 'Total phenols', 'Flavanoids', 'Nonflavanoid phenols', 'Proanthocyanins', 'Color intensity', 'Hue', 'OD280/OD315 of diluted wines', 'Proline')
wine_data <- read.table(file="/Users/19892/OneDrive/Documents/STATS506/Data/wine.data", sep=",", header=TRUE, col.names = col_names)

wine_data
```

### Part B: The data contains three different classes of wine (1, 2, 3) as seen in the above data frame. 

### Part C: Next, I will use the data to answer the following questions:

1.  What class does the highest alcohol content belong to?

    We will do this using the 'which.max()' function in the Alcohol column from 'wine_data'.

    ```{r}
    max_alcohol = wine_data[which.max(wine_data$Alcohol),]
    max_alcohol
    ```

Here, we find that the highest alcohol content belongs to class 1, being 14.83 ABV (alcohol by volume).

2.   The wine with the lowest alcohol content belongs to which class?

    For this, we will use a similar function, except with the 'which.min()' function in the Alcohol column from 'wine_data'.

    ```{r}
    min_alcohol = wine_data[which.min(wine_data$Alcohol),]
    min_alcohol
    ```

Here, we find that the lowest alcohol content belongs to class 2, being 11.03 ABV.

3.   German beers have, on average 114 mg/l of magnesium. How many of these wines have higher levels of magnesium than that? We will assume that the magnesium reported in the data is in the same units.
